PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-06-20 ijcis-pos-model/sources/timeRepresentation.tex
final draft

\subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
%The datatypes for an  $X$ are shown in Table \ref{}.
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:time-point-types}Values for the time point data type.}
\centerline{\small TIME POINT DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
Subtype & Value & Representation \\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t,\forall t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t,\forall t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
}

@@

%\begin{definition}
\textbf{Ill-known time interval}\\
An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known. The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$. The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
Consider an ill-known value $T$, a binary relationship $B_r$ one of $\left(\leq, <, >, \geq\right)$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.

The possibility and necessity measures are defined by:
\begin{eqnarray}
\label{eq:open-pos-nec}
I_p \left(C \right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
I_n \left(C \right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
\end{eqnarray}
Where the values for the binary relations $\Rp$ and $\Rn$ are shown in Table \ref{tbl:open-pos-nec-rels}.
\end{definition}

\vglue13pt
%\begin{table}[htbp]
\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
\centerline{\small Relations}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c}\\
\hline
Constraint & $\Rp$ & $\Rn$\\ \hline
$C\stackrel{\triangle}{=}\left(<,T\right)$ & $>$ & $\leq$ \\
$C\stackrel{\triangle}{=}\left(\leq,T\right)$ & $\geq$ & $<$ \\
$C\stackrel{\triangle}{=}\left(>,T\right)$ & $<$ & $\geq$ \\
$C\stackrel{\triangle}{=}\left(\geq,T\right)$ & $\leq$ & $>$ \\
\hline\\
\end{tabular}
\label{tbl:open-pos-nec-rels}
}

@@

\begin{example}
\label{ex:open-time-interval}
Consider an ill-known time interval given by $\left[\FB, Y \right]$. Consider also that in this case, $Y= \left[15/10/2012,3,4 \right] $. Figure \ref{fig:example-open-time-interval} shows the representation for $Y$. The user wants to obtain the possibility and the necessity measures for the $\FB$ part of the interval.
\end{example}

@@

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/open-time-interval.pdf}
\label{fig:example-open-time-interval}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
\vspace*{13pt}

@@

\subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
In order to properly represent  an ill-known time interval in a database, some datatypes are needed. Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed. Table \ref{tbl:time-interval-types} shows all the possible datatypes combinations for the  for an ill-known time interval denoted by $\left[X, Y\right]$.

\vglue13pt
%\begin{table}[htbp]
\tcap{All the possible combination of values for the time interval $\left[X, Y\right]$. The subtypes refer to Table \ref{tbl:time-point-types}.}
\centerline{\small TIME INTERVAL DATA TYPE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{p{0.15\columnwidth} p{0.15\columnwidth} p{0.5\columnwidth}}\\
\hline
Subtype for $X$  & Subtype for $Y$  & Description\\
\hline
1 or 2 & 1 or 2 &  An ill-known time interval.\\
3 & 3 & An unknown time interval. \\
$\FB$ & 1 or 2 & A left-open time interval.\\
1 or 2 & $\UC$ & A right-open time interval.\\
\hline\\
\end{tabular}
\label{tbl:time-interval-types}
}

>>>>

\subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
%The datatypes for an  $X$ are shown in Table \ref{}.
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:time-point-types}Values for the time point data type.}
\centerline{\small DIFFERENT VALUES FOR A TIME POINT}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
Subtype & Value & Representation \\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t,\forall t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t,\forall t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
}
\end{samepage}

@@

An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known. %The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$. The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.

\begin{definition}
\emph{Ill-known time interval}
Let $\T$ be the time domain, and $X$, $Y$ two ill-known values in the time domain. An ill-known time interval is given by $\left[X, Y\right]$. The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}. We will note $\I_{PVP}$ the set of all the ill-known time intervals.

\end{definition}

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.

The possibility and necessity measures are defined by:
\begin{eqnarray}
\label{eq:open-pos-nec}
I_p \left(C \right) &=& \Big(\sup_{r\in\T, r \Rp\  w}\pi_{T}(w)\Big)\\
I_n \left(C \right) &=& \Big(\inf_{r\in\T, r \Rn\  w}1-\pi_{T}(w)\Big)
\end{eqnarray}
Where the values for the binary relations $\Rp$ and $\Rn$ are shown in Table \ref{tbl:open-pos-nec-rels}.
\end{definition}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
\centerline{\small Relations}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c}\\
\hline
Constraint & $\Rp$ & $\Rn$\\ \hline
$C\stackrel{\triangle}{=}\left(<,T\right)$ & $>$ & $\leq$ \\
$C\stackrel{\triangle}{=}\left(\leq,T\right)$ & $\geq$ & $<$ \\
$C\stackrel{\triangle}{=}\left(>,T\right)$ & $<$ & $\geq$ \\
$C\stackrel{\triangle}{=}\left(\geq,T\right)$ & $\leq$ & $>$ \\
\hline\\
\end{tabular}
\label{tbl:open-pos-nec-rels}
}
\end{samepage}

@@

\begin{example}
\label{ex:open-time-interval}
Consider an ill-known time interval given by $\left[\FB, Y \right]$. Consider also that, in this case, $Y= \left[15/10/2012,3,4 \right] $. Figure \ref{fig:example-open-time-interval} shows the representation for $Y$. The user wants to obtain the possibility and the necessity measures for the $\FB$ part of the interval.
\end{example}

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/open-time-interval.pdf}
\label{fig:example-open-time-interval}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
\vspace*{13pt}
\end{samepage}

@@

\subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
In order to properly represent  an ill-known time interval in a database, some datatypes are needed. Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed. Table \ref{tbl:time-interval-types} shows all the possible values that can be used to represent an ill-known time interval denoted by $\left[X, Y\right]$.

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{All the possible combination of values for the time interval $\left[X, Y\right]$. The subtypes refer to Table \ref{tbl:time-point-types}.}
\centerline{\small TIME INTERVAL DATA TYPE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{p{0.15\columnwidth} p{0.15\columnwidth} p{0.5\columnwidth}}\\
\hline
Subtype for $X$  & Subtype for $Y$  & Description\\
\hline
1 or 2 & 1 or 2 &  An ill-known time interval.\\
3 & 3 & An unknown time interval. \\
$\FB$ & 1 or 2 & A left-open time interval.\\
1 or 2 & $\UC$ & A right-open time interval.\\
\hline\\
\end{tabular}
\label{tbl:time-interval-types}
}
\end{samepage}
